Distributed Web Mining of Ethereum\thanks{Supported by organization x.}

### 2018-04-02 paper/paper.tex
changed paper and figs

\subsection{Our Contributions}
Our research involves providing implementations of a distributed, web browser based Ethereum miners in \verb|JavaScript| and \verb|WebAssembly|, and studying whether or not it is feasible for real world purposes. While there has been prior academic work done in related fields, none of the distributed web browser implementations in these works do not involve crypto-currency mining. While CoinHive does perform distributed mining using web browsers, it is closed source, and also works with a different crypto-currency (Monero). Hence, we are pioneering in offering an open source architecture for Ethereum mining with web browsers. Our miner gives us a hashrate of about 39kH/s, which, given the flexibility of the Ethereum framework, is useful and practical for many applications like web authentication rate limiting, website usage tracking for advertisement companies, private Ethereum networks, web content monetization, as discussed below.

@@

Another potential application could involve web-authentication rate limiting. Given the multitude of tools available brute forcing web login pages ~\cite{hydra} ~\cite{burpsuite}, it is painfully easy to brute force authentication pages these days. Currently, the way website owners mitigate this issue is by locking out a user for a certain amount of time after a fixed number of unsuccessful login attempts. 

@@

Finally, we would also like to note that Ethereum is an extremely flexible currency in the sense that it allows for private coin networks -- i.e., networks that do not on mine the actual, public Ethereum block-chain, but rather on private (and often smaller) instances of the crypto-currency. Our browser miner can be used on any such private network to serve the network owner whatever he/she wishes to implement with the mining PoW.

\section{Background}
% \subsection{Ethereum}
% \trishita{Not sure if I need this paragraph, since I think the readers will probably be familiar with what crypto-curriencies are}
\subsection{JavaScript and WebAssembly}
For very long, \verb|JavaScript| has been the only language available to make dynamic client side web content. However, the situation has changed since the advent of \verb|WebAssembly| in 2016. Per its creators, ``\verb|WebAssembly| is a binary instruction format for a stack-based virtual machine. Wasm is designed as a portable target for compilation of high-level languages like \verb|C/C++/Rust|, enabling deployment on the web for client and server applications. The \verb|WebAssembly| stack machine is designed to be encoded in a size- and load-time-efficient binary format. \verb|WebAssembly| aims to execute at native speed by taking advantage of common hardware capabilities available on a wide range of platforms'' ~\cite{webAssembly}. The language is meant to improve performance for computationally intensive use cases such as image/video editing, games, streaming, etc ~\cite{webAssembly}. This, and given the fact that CoinHive is also implemented in \verb|WebAssembly| ~\cite{coinhive}, makes the language a natural choice for us to implement a miner in. In fact, as of now, \verb|WebAssembly| is supported by 4 major browser versions -- Firefox, Safari, IE, and Chrome ~\cite{webAssembly}, which makes our miner implementation extremely versatile.

@@

Mining is performed by starting with the current block header hash and nonce, which are combined to get a 128B wide ``Mix'' (another byte array). The Mix is then used to fetch a specific slice of the DAG from memory. (Each slice is 512 bits long, or 16 4 byte integers) After this, the Mix is updated with the fetched part of the DAG. Then, this updated Mix is used to fetch a new part of the DAG. This process of sequentially fetching parts of the DAG is repeated 64 times, and the final value of the Mix is put through a transformation to obtain a 32 byte digest. This digest is then compared to the threshold, and if it is lesser than it, the nonce is valid and the block is successfully mined and can be broadcast to the network. However, if the digest is greater than the threshold, the nonce was unsuccessful, and the entire process must be repeated with a new nonce ~\cite{Ethmining}. This procedure for mining is depicted in Figure~\ref{fig:ethash}.

It is important to note that the parts of the DAG that are used to compute the hash for a particular block depend on the nonce used, hence no way to pre-determine which parts will be useful to have in memory. Hence this forces miners to store entire DAG in memory, making mining ``Memory Hard''. This is intended to even the playing field, as ASIC miners no longer have any advantage as they do for Bitcoin. GPU miners, however, do have an advantage, normally having hashrates of over 10 times that of a CPU miner.
While mining is memory intensive, verification is relatively lightweight. This is because of the property that each ``slice'' in the DAG depends on a set of pseudo-randomly selected items from the Cache. Hence, the Cache is used to regenerate only the specific slices of the DAG that are needed to recalculate the hash for the particular nonce. And so only the Cache needs to be stored by clients that just perform verification. We use this property of being able to generate the parts of the DAG as needed to our advantage in order to alleviate some of the memory and network bandwidth restrictions that browsers typically face.

As time goes on, mining Ethereum becomes more and more difficult, as the size of the DAG and Cache used for the Ethereum increases every Epoch. Currently the DAG is more than 2GB in size, causing many GPU miners with cards of 2GB of memory to suffer in terms of hashrate. 
% The Proof-of-Work difficulty and total hashrate of the network has been steadily increasing over time, with the current difficulty at 1612.737TH and the respective hashrate at 109.138TH/s ~\cite{etherscan}. This reflects the design of mining one block approximately every 12 seconds. 
% As for mining rewards, the reward for mining a block has gone from 5 Ether to 3 ether, though the difficulty has significantly decreased, both due to the Byzantium fork.

For a normal Ethereum miner to function, the current block headers and the DAG are required to even begin mining. As we simply cannot afford to transfer the entire DAG to the browser and store it in memory, we propose an an alternative approach addresses some of these concerns.

\begin{figure}[h]
\centering
\includegraphics[width=300px,keepaspectratio]{ethash_algorithm.png}
\caption{\label{fig:ethash} Ethash}
\end{figure}

\subsection{Challenges}
Given that we now know that the crux of having a good hashrate for Ethereum lies within having quick access to DAG, one of the biggest challenges with mining it on a distributed ensemble of browsers would be to transfer and store this data structure within each worker browser. However, given that the DAG for an Ethereum network is at least 1GB (and growing) in size transferring it to the browser every time a client loads our web page would be impractical. Furthermore, even if we somehow manage to transfer the DAG to the client when the web page loads, the browser would still have to store it in memory, thereby using a sizeable amount of resources on the client machine. We attempt to address some of these concerns through our approach as discussed in the next section.

\section{Our Approach}
We propose a lazy-evaluation based approach to alleviate the network and memory requirements for mining ethereum in a distributed scenario on browsers. Specifically, as soon as each browser connects to the webserver, the server sends it the current header of the block being mined and the Cache. Once the browser receives the Cache, it allocates an array buffer to store the DAG slices that fall between indices 0 and $n$ that will be computed in the future. (Note that the $n$ is a variable parameter reflecting the length of the buffer with which we experiment in our analysis). Once the buffer is allocated, the browser can start iterating over nonces to compute hashes.

@@

Our architecture was centered around two client-side Ethereum miners written in \verb|JavaScript| and \verb|WebAssembly|. The central node that co-ordinates all workers (browsers) was an improvised version of \verb|geth| ~\cite{geth}, a real world Ethereum miner written in \verb|Go|. \verb|Geth| typically runs as a standalone miner that mines on the machine it is running. We modified it so that instead of mining all by itself, it simply sends over the necessary data needed to mine (the hash of the Block Header and the Cache) to any client that connects to it on port \verb|9000|. After recieving the necessary data, the browser allocates a buffer for the DAG in order to store future slices. Note that the buffer for the DAG is implemented an array of ints (with a DAG slice being 16 ints long), so as to make each lookup in the buffer $\theta(1)$. At this point, it can then begin to search for a solution.

For our miner, we modeled our \verb|JavaScript| implementation based on the ~\verb|node.js| implementation of Ethash ~\cite{ethash}. To begin, the miner creates a random nonce and the computes the hash (using the Cache and the buffered DAG) as discussed in the previous section. It will continue to perform this action on new nonces until one of two scenarios occur; it finds a nonce such that the computed hash is below the given threshold or the process has timed out. In the former case, the browser submits the result back to the node and then asks the node for a new block header and the Cache. Otherwise if it has timed out without finding a result, the browser simply polls the node for the current block header and Cache . This process will continue until the user moves away from the website or closes their browser. 
We must point out that this time out is necessary since we want the browser to work with the most recent block header and Cache. The block header can become stale if that particular block has already been mined, and the Cache can become stale if the Ethereum network transitions into a new Epoch (happens once every 30,000 blocks). 
We must point out that both our current implementations in \verb|JavaScript| and \verb|WebAssembly| require no external dependencies, and therefore can be directly embedded into any website. For a diagramatic view our our design, see Figures~\ref{fig:hybridArchitecture}.

@@

\subsection{Theoretical Analysis}
With this lazy-evaluation based approach, we can calculate exactly how many hashes we would have to compute in order to fill up a buffer of a given size. This is important because once the buffer in the browser fills up, the hash rate that the browser gives becomes steady. Hence, calculating the expected number of hashes it takes to fill the buffer allows us to determine how much time will the browser take to reach a steady state hash rate. This approach is a variation of the coupon collector problem \cite{couponCollector}, and has been described in the following theorem:
\trishita{To Do: change these expressions to upper and lower bounds since we are dealing with a truncated summation}
\begin{theorem}[Expectation of Number of Hashes]
\label{expectedHashes}\\
\makebox[\textwidth][l]{If we define}
\makebox[\textwidth][c]{Number of DAG slices = n}
\makebox[\textwidth][c]{Size of buffer as a fraction of the total DAG = $f_1$; $f_1$=[0,1]}
\makebox[\textwidth][c]{Expected number of hashes it takes to fill the buffer = E(H)} \\
Then the expected number of hashes $E(H)$ it takes to fill up a buffer of fraction $f_1$ of a DAG with $n$ slices is given by:
\begin{gather}
E(H) = \frac{n\log{nf_1}}{64}
\end{gather}
\end{theorem}
\begin{proof}
Let us assume we have a buffer for our DAG that is $f_1$ as large as the entire DAG -- i.e., the buffer can only store dag slices with indices $i$ such that $i \in [0,nf_1]$. Hence, each time we calculate a DAG slice from the Cache, we may or may not buffer it depending on whether it falls within the index rage. Now, let us assume that  we have buffered $x$ slices so far, the probability of obtaining a new slice is given by:
\begin{gather}
	P_x = \frac{nf_1-x}{n}
\end{gather}
The event of obtaining a new slice after having already collected $x$ out of the $nf_1$ slices that can be buffered can be modeled as a geometric distribution. Hence, the expected number of trials $t_x$ it takes to get a new  slice after having seen $x$ out of the $nf_1$ slices is:
\begin{gather}
	E(t_x) = \frac{1}{P_x} = \frac{n}{nf_1-x}
\end{gather}
By linearity of expectation, the total number of trials it would take to see all $x$ slices is:
\begin{gather}
	E(t) = \sum_{x=0}^{nf_1-1}E(t_x) = \sum_{x=0}^{nf_1-1}P_x = \sum_{x=0}^{nf_1-1}\frac{n}{nf_1-x} = n\sum_{x=0}^{nf_1-1}\frac{1}{nf_1-x}
\end{gather}
Since the summation represents the harmonic series, the expected value can be approximated as follows:
\begin{gather}
	E(t) = n\log{nf_1}
\end{gather}
Equation ~\ref{eq:7} is in terms of the number of individual accesses of the DAG. In order to get the expected number of Hashes (as opposed to the expected number of individual DAG slice accesses), we need to divide by 64 since each hash involves 64 accesses of the DAG. Hence: 
\begin{gather}
	E(H) = \frac{E(t)}{64} = \frac{n\log{nf_1}}{64}
\end{gather}
\end{proof}
Now, let us do a simple calculation for one extreme case, where we allocate a buffer as large as the entire DAG on the browser (in order to get an upper bound on the number of hashes it would take to fill this buffer up). Here $f_1 = 1$ and $n = 16777186$ (for our private test network), which, using ~\ref{eq:6}, gives us $E(H) = 4.36*10^6$ hashes. Hence, it takes around 4 million hashes to fill up a buffer as large as the entire DAG. If $f_1 < 1$, the number of hashes should be less than 4 million. So effectively, 4 million hashes is a very lose upper bound on the number of hashes needed to fill up a DAG buffer. However, since 4 million hashes can be a lot for a browser, we will now demonstrate how we can get away with  better

Now let us say that we do not want to wait for the entire buffer to fill up -- how many hashes would we have to compute until a reasonable fraction of the buffer is full? For this, we have the following corollary:
\begin{corollary}
\label{earlyTermination}
\makebox[\textwidth][l]{If we define,}
\makebox[\textwidth][c]{Fraction of DAG buffer to be filled = $f_2$; $f_2$ $\in$ [0,1]}
\makebox[\textwidth][l]{Then,}
    \begin{gather}
		E(H) = \frac{n}{64}\log{\frac{1}{1-f_2-\frac{1}{nf_1}}}
	\end{gather}
\end{corollary}
\begin{proof}
Then, for a buffer with space for $f_1$ slices, the number of slices we want to fill it with in order for a fraction $f_2$ of it to be full would be: \\
\makebox[\textwidth][c]{Number of slices to fill DAG buffer with = $nf_2f_1$}
This would modify equation \ref{eq:6} for the expected number of trials to fill up a fraction $f_2$ of the DAG buffer to be as follows:
\begin{gather}
    E(t) = \sum_{x=0}^{nf_2f_1}P_x = \sum_{x=0}^{nf_2f_1}\frac{n}{nf_1-x} = n\sum_{x=0}^{nf_2f_1}\frac{1}{nf_1-x}
\end{gather}
This is essentially an early termination of the summation in equation ~\ref{eq:6}, and can be expressed as follows:
\begin{gather}
    E(t) = n\sum_{x=0}^{nf_2f_1}\frac{1}{nf_1-x} = n(\sum_{x=0}^{nf_1-1}\frac{1}{nf_1-x} - \sum_{x=nf_2f_1+1}^{nf_1-1}\frac{1}{nf_1-x})\\
    E(t) = n(\log{nf_1} - \log{(nf_1-nf_2f_1-1)}) \\
    E(t) = n\log{\frac{1}{1-f_2-\frac{1}{nf_1}}}
\end{gather}
Again, we divide by 64 to obtain the expected number of hash computations:
\begin{gather}
    E(H) = \frac{n}{64}\log{\frac{1}{1-f_2-\frac{1}{nf_1}}}
\end{gather}
\end{proof}
Now, using ~\ref{eq:13}, if we set $n = 16777186$, $f_1 = 1$ and $f_2 = 0.95$, we get $E(H) = 785312$ hashes

@@

Hence, we see that even computing merely 800,000 hashes fills a DAG buffer as large as the entire DAG by more than 95\%, as opposed to calculating 4 million to fill the buffer entirely. (Note that this number should be smaller for smaller buffer sizes; so this is again a rough upper bound). This suggests that the fuller the buffer is, the marginal cost of computing more hashes outweighs the rate at which the buffer is filled up. This also means that reaching a reasonably steady state is not as hard as it seems at face value, making this approach viable. Given this result, we compute 800,000 hashes for our experiments.
% Beyond that, the marginal cost of computing more hashes outweighs the rate at which the DAG is filled up. Hence, as we will see in the next section, for our experiments, we compute 800,000 hashes.

\section{Results}
\subsection{Experimental set up}
Our experimental set up consisted of a machine with an Intel i7-7700HQ processor with 8 cores and 16GB ram. These results were obtained from a private Ethereum test network at Epoch 0. The DAG size was 16777186 slices (1.074GB). The cache size was 1.677MB. We ran the implementations in ~\verb|JavaScript| and ~\verb|WebAssembly| in the browser, and a native miner written in ~\verb|C++| that employs the same approach outside of the browser for control results. Each miner was till 800,000 hashes were computed and the hash rate and buffer hit rate were sampled every 10,000 hashes. (Note that both the hash rate and buffer hit rate sampled at a particular time reflected the values over the 10,000 most recent hashes).

>>>>

\subsection{Our Contributions}
Our research involves providing implementations of a distributed, web browser based Ethereum miners in \verb|JavaScript| and \verb|WebAssembly|, and studying whether or not it is feasible for real world purposes. While there has been prior academic work done in related fields, none of the distributed web browser implementations in these works involve crypto-currency mining. While CoinHive does perform distributed mining using web browsers, it is closed source, and also works with a different crypto-currency (Monero). Hence, we are pioneers in offering an open source architecture for Ethereum mining with web browsers. Our miner gives us a hashrate of about 39kH/s, which, given the flexibility of the Ethereum framework, is useful and practical for many applications like web authentication rate limiting, website usage tracking for advertisement companies, private Ethereum networks, web content monetization, as discussed below.

@@

Another potential application could involve web-authentication rate limiting. Given the multitude of tools available brute forcing web login pages ~\cite{hydra} ~\cite{burpsuite}, it is extremely easy to brute force authentication pages these days. Currently, the way website owners mitigate this issue is by locking out a user for a certain amount of time after a fixed number of unsuccessful login attempts. 

@@

Finally, we would also like to note that Ethereum is an extremely flexible currency in the sense that it allows for private coin networks -- i.e., networks that do not mine on the actual, public Ethereum block-chain, but rather on private (and often smaller) instances of the crypto-currency. Our browser miner can be used on any such private network to serve the network owner whatever he/she wishes to implement with the mining PoW.

\section{Background}
% \subsection{Ethereum}
% \trishita{Not sure if I need this paragraph, since I think the readers will probably be familiar with what crypto-curriencies are}
\subsection{JavaScript and WebAssembly}
For very long, \verb|JavaScript| has been the only language available to make dynamic client side web content. However, the situation has changed since the advent of \verb|WebAssembly| in 2016. As per its creators, ``\verb|WebAssembly| is a binary instruction format for a stack-based virtual machine. Wasm is designed as a portable target for compilation of high-level languages like \verb|C/C++/Rust|, enabling deployment on the web for client and server applications. The \verb|WebAssembly| stack machine is designed to be encoded in a size- and load-time-efficient binary format. \verb|WebAssembly| aims to execute at native speed by taking advantage of common hardware capabilities available on a wide range of platforms'' ~\cite{webAssembly}. The language is meant to improve performance for computationally intensive use cases such as image/video editing, games, streaming, etc ~\cite{webAssembly}. This, and given the fact that CoinHive is also implemented in \verb|WebAssembly| ~\cite{coinhive}, makes the language a natural choice for us to implement a miner in. In fact, as of now, \verb|WebAssembly| is supported by 4 major browser versions -- Firefox, Safari, IE, and Chrome ~\cite{webAssembly}, which makes our miner implementation extremely versatile.

@@

Mining is performed by starting with the current block header hash and nonce, which are combined to get a 128B wide ``Mix'' (another byte array). The Mix is then used to fetch a specific page of the DAG from memory. (Each page of the DAG 2 slices long, and each slice in turn is 64 Bytes long) After this, the Mix is updated with the fetched part of the DAG. Then, this updated Mix is used to fetch a new part of the DAG. This process of sequentially fetching parts of the DAG is repeated 64 times, and the final value of the Mix is put through a transformation to obtain a 32 byte digest. This digest is then compared to the threshold, and if it is lesser than it, the nonce is valid and the block is successfully mined and can be broadcast to the network. However, if the digest is greater than the threshold, the nonce is unsuccessful, and the entire process must be repeated with a new nonce ~\cite{Ethmining}. This procedure for mining is depicted in Figure~\ref{fig:ethash}.

It is important to note that the parts of the DAG that are used to compute the hash for a particular block depend on the nonce used, hence there is no way to pre-determine which parts will be useful to have in memory. This therefore forces miners to store entire DAG in memory, making mining ``Memory Hard''. This is intended to even the playing field, as ASIC miners no longer have any advantage as they do for Bitcoin. GPU miners, however, do have an advantage, normally having hashrates of over 10 times that of a CPU miner.
While mining is memory intensive, verification is relatively lightweight. This is because of the property that each ``slice'' in the DAG depends on a set of pseudo-randomly selected items from the Cache. Hence, the Cache is used to regenerate only the specific slices of the DAG that are needed to recalculate the hash for the particular nonce. And so only the Cache needs to be stored by clients that perform just verification. We use this property of being able to generate the parts of the DAG as needed to our advantage in order to alleviate some of the memory and network bandwidth restrictions that browsers typically face.

As time goes on, mining Ethereum becomes more and more difficult, as the size of the DAG and Cache increases with every Epoch. Currently the DAG is more than 2GB in size, causing many GPU miners with cards of 2GB of memory to suffer in terms of hashrate. 
% The Proof-of-Work difficulty and total hashrate of the network has been steadily increasing over time, with the current difficulty at 1612.737TH and the respective hashrate at 109.138TH/s ~\cite{etherscan}. This reflects the design of mining one block approximately every 12 seconds. 
% As for mining rewards, the reward for mining a block has gone from 5 Ether to 3 ether, though the difficulty has significantly decreased, both due to the Byzantium fork.

For a normal Ethereum miner to function, the current block headers and the DAG are required to even begin mining. As we simply cannot afford to transfer the entire DAG to the browser and store it in memory, we propose an an alternative approach to address some of these concerns.

\begin{figure}[h]
\centering
\includegraphics[width=300px,keepaspectratio]{Ethash.pdf}
\caption{\label{fig:ethash} Ethash}
\end{figure}

\subsection{Challenges}
\trishita{Too long of a sentence; need to reword}Given that we now know that the crux of having a good hashrate for Ethereum lies in having quick access to the DAG, one of the biggest challenges with mining it on a distributed ensemble of browsers would be to transfer and store this data structure within each worker browser. However, given that the DAG for an Ethereum network is at least 1GB (and growing) in size transferring it to the browser every time a client loads our web page would be impractical. Furthermore, even if we somehow manage to transfer the DAG to the client when the web page loads, the browser would still have to store it in memory, thereby using a sizeable amount of resources on the client machine. We attempt to address some of these concerns through our approach as discussed in the next section.

\section{Our Approach}
We propose a lazy-evaluation based approach to alleviate the network and memory requirements for mining Ethereum in a distributed scenario on browsers. Specifically, as soon as each browser connects to the webserver, the server sends it the current header of the block being mined and the Cache. Once the browser receives the Cache, it allocates an array buffer to store the DAG slices that fall between indices 0 and $n$ that will be computed in the future. (Note that the $n$ is a variable parameter reflecting the length of the buffer with which we experiment in our analysis). Once the buffer is allocated, the browser can start iterating over nonces to compute hashes.

@@

Our architecture was centered around two client-side Ethereum miners written in \verb|JavaScript| and \verb|WebAssembly|. The central node that co-ordinates all workers (browsers) was an improvised version of \verb|geth| ~\cite{geth}, a real world Ethereum miner written in \verb|Go|. \verb|Geth| typically runs as a standalone miner that mines on the machine it is running. We modified it so that instead of mining all by itself, it simply sends over the necessary data needed to mine (the hash of the Block Header and the Cache) to any client that connects to it on port \verb|9000|. After recieving the necessary data, the browser allocates a buffer for the DAG in order to store future slices. Note that the buffer for the DAG is implemented as an array of ints (with a DAG slice being 16 ints long), so as to make each lookup in the buffer $\theta(1)$. At this point, it can then begin to search for a solution.

For our miner, we modeled our \verb|JavaScript| implementation based on the ~\verb|node.js| implementation of Ethash ~\cite{ethash}. To begin, the miner creates a random nonce and computes the hash (using the Cache and the buffered DAG) as discussed in the previous section. It will continue to perform this action on new nonces until one of two scenarios occur; it finds a nonce such that the computed hash is below the given threshold or the process has timed out. In the former case, the browser submits the result back to the node and then asks the node for a new block header and the Cache. Otherwise if it has timed out without finding a result, the browser simply polls the node for the current block header and Cache . This process will continue until the user moves away from the website or closes their browser. 
We must point out that this time out is necessary since we want the browser to work with the most recent block header and Cache. The block header can become stale if that particular block has already been mined, and the Cache can become stale if the Ethereum network transitions into a new Epoch (happens once every 30,000 blocks). 
We must point out that both our current implementations in \verb|JavaScript| and \verb|WebAssembly| require no external dependencies, and therefore can be directly embedded into any website. For a diagramatic view our our design, see Figures~\ref{fig:hybridArchitecture}.

@@

\subsection{Theoretical Analysis}
With this lazy-evaluation based approach, we can calculate exactly how many hashes we would have to compute in order to fill up a buffer of a given size. This is important because once the buffer in the browser fills up, the hash rate that the browser gives is at its maximum, steady-state. In this section, we will mathematically demonstrate that filling a buffer the size of the DAG up should take at most 4.5 million hashes, while filling up 95\% of this buffer takes merely 800k hashes. This is a significant difference since in our experimental set-up, the browser takes X s to compute 4.5 million hashes, while only Y s to compute 800k hashes. The approach we take is similar to the Coupon Collector problem \cite{couponCollector}, and is discussed below:
% \begin{theorem}[Expectation of Number of Hashes]
% \label{expectedHashes}\\
\makebox[\textwidth][l]{If we define}
\makebox[\textwidth][c]{Number of DAG slices = n}
\makebox[\textwidth][c]{Size of buffer as a fraction of the total DAG = $f_1$; $f_1$=[0,1]}
\makebox[\textwidth][c]{Expected number of hashes it takes to fill the buffer = E(H)}
\makebox[\textwidth][c]{nth Harmonic Number = $H_n$}
\makebox[\textwidth][c]{Euler-Mascheroni constant = $\gamma = 0.57721566...$}
\makebox[\textwidth][c]{Probability of obtaining a new slice after having buffered x slices = $P_x = \frac{nf_1-x}{n}$}
\makebox[\textwidth][c]{Expected number of trials $t_x$ to get a new slice after having seen} 
\makebox[\textwidth][c]{$x$ out of $nf_1$ slices = 
$E(t_x) = \frac{1}{P_x} = \frac{n}{nf_1-x}$}
\makebox[\textwidth][c]{Expected number of trials it would take to see all $x$ slices = $E(t)$}
Note that for simplification, we assume that the DAG is randomly sampled in order to compute hashes. In reality, this is not entirely true and hence the results we obtain from this analysis are more conservative than what we see in practice. In fact, while we show here that filling up 95\% of a buffer as big as the DAG takes 800k hashes, whereas in reality it only takes half as many hashes~\ref{section:Results} \\
% ~\trishita{add reference to results}
From the Coupon Collector Problem~\cite{couponCollector}, it follows that:
\begin{gather}
E(t) = \sum_{x=0}^{nf_1-1}E(t_x) = \sum_{x=0}^{nf_1-1}\frac{1}{P_x} = n\sum_{x=0}^{nf_1-1}\frac{1}{nf_1-x} = nH_{nf_1}
\end{gather}

It then follows from bounds on the nth Harmonic Number~\cite{harmonicNumber} that the expected number of hashes $E(H)$ it takes to fill up a buffer of fraction $f_1$ of a DAG with $n$ slices is bounded by:
\begin{gather}
\frac{n}{64}[\ln{nf_1} + \frac{1}{2nf_1} + \gamma - \frac{1}{12({nf_1})^2+2\frac{7-12\gamma}{2\gamma - 1}}] <= E(H) \\ E(H)< \frac{n}{64}[\ln{nf_1} + \frac{1}{2nf_1} + \gamma - \frac{1}{12({nf_1})^2+\frac{6}{5}}] 
\end{gather}

% this was the proof for the expression above
% % \end{theorem}
% % \begin{proof}
% Let us assume we have a buffer for our DAG that is $f_1$ as large as the entire DAG -- i.e., the buffer can only store dag slices with indices $i$ such that $i \in [0,nf_1]$. Hence, each time we calculate a DAG slice from the Cache, we may or may not buffer it depending on whether it falls within the index rage. Now, let us assume that  we have buffered $x$ slices so far, the probability of obtaining a new slice is given by:
% \begin{gather}
% 	P_x = \frac{nf_1-x}{n}
% \end{gather}
% The event of obtaining a new slice after having already collected $x$ out of the $nf_1$ slices that can be buffered can be modeled as a geometric distribution. Hence, the expected number of trials $t_x$ it takes to get a new  slice after having seen $x$ out of the $nf_1$ slices is:
% \begin{gather}
% 	E(t_x) = \frac{1}{P_x} = \frac{n}{nf_1-x}
% \end{gather}
% By linearity of expectation, the total number of trials it would take to see all $x$ slices is:
% \begin{gather}
% 	E(t) = \sum_{x=0}^{nf_1-1}E(t_x) = \sum_{x=0}^{nf_1-1}P_x = \sum_{x=0}^{nf_1-1}\frac{n}{nf_1-x} = n\sum_{x=0}^{nf_1-1}\frac{1}{nf_1-x}
% \end{gather}
% Since the summation represents the harmonic series, the expected value can be approximated as follows:
% \begin{gather}
% 	E(t) = n\log{nf_1}
% \end{gather}
% Equation ~\ref{eq:7} is in terms of the number of individual accesses of the DAG. In order to get the expected number of Hashes (as opposed to the expected number of individual DAG slice accesses), we need to divide by 64 since each hash involves 64 accesses of the DAG. Hence: 
% \begin{gather}
% 	E(H) = \frac{E(t)}{64} = \frac{n\log{nf_1}}{64}
% \end{gather}
% \end{proof}

The factor of 64 comes from the fact that calculating each hash requires sampling the DAG 64 times. Given this, let us do a simple calculation for one extreme case, where we allocate a buffer as large as the entire DAG on the browser (in order to get an upper bound on the number of hashes it would take to fill this buffer up). Here $f_1 = 1$ and $n = 16777186$ (for our private test network), which, using ~\ref{eq:6}, gives us $4512210 < E(H) <= 4512210$ hashes. Hence, it takes around 4.5 million hashes to fill up a buffer as large as the entire DAG. If $f_1 < 1$, the number of hashes should be less than 4.5 million, thus making it a very loose upper bound on the number of hashes needed to fill up a DAG buffer (and thereby reaching a steady-state hash rate). 
% \trishita{To Do: reword the next two sentences to motivate the analysis that follows}
In practice, however, we do not need to wait till the buffer is \emph{completely} full till we see good hash rates. We will demonstrate how we can get away with a much smaller number of hashes to fill up \emph{almost all} of the buffer, thereby reaching very good hash rates relatively quickly.
% lower bound (python exp): (n/64.0)*(log(n) + 1.0/(2*n) + g - 1.0/(12*(n**2) + (2*(7-12*g)/(2*g-1))))
% upper bound (python exp): (n/64.0)*(log(n) + 1.0/(2*n) + g - 1.0/(12*(n**2) + 6.0/5))

% Now let us say that we do not want to wait for the entire buffer to fill up -- how many hashes would we have to compute until a reasonable fraction of the buffer is full? For this, we have the following corollary:
% \begin{corollary}
% \label{earlyTermination}
\makebox[\textwidth][l]{If we further define,}
\makebox[\textwidth][c]{Fraction of DAG buffer to be filled = $f_2$; $f_2$ $\in$ [0,1]}
\makebox[\textwidth][c]{$x  = E(H) - \frac{n}{64}\ln{\frac{nf_1}{nf_1 - nf_2f_1 - 1}} - \frac{1}{128f_1} + \frac{n}{768({nf_1})^2+128\frac{7-12\gamma}{2\gamma - 1}}$}
\makebox[\textwidth][l]{Then,}
    \begin{gather}
   \frac{n}{768({nf_1 - nf_2f_1 - 1})^2+128\frac{7-12\gamma}{2\gamma - 1}} - \frac{1}{768({nf_1})^2+128\frac{7-12\gamma}{2\gamma - 1}} < x \\
   x <= \frac{n}{768({nf_1-nf_2f_1-1})^2+\frac{6}{5}} - \frac{n}{12({nf_1})^128+\frac{6}{5}} 
	\end{gather}
% \end{corollary}
% \begin{proof}
Then, for a buffer with space for $f_1$ slices, the number of slices we want to fill it with in order for a fraction $f_2$ of it to be full would be: \\
\makebox[\textwidth][c]{Number of slices to fill DAG buffer with = $nf_2f_1$}
This would modify equation \ref{eq:6} for the expected number of trials as follows:
\begin{gather}
    E(t) = \sum_{x=0}^{nf_2f_1}\frac{1}{P_x} = \sum_{x=0}^{nf_2f_1}\frac{n}{nf_1-x} = n\sum_{x=0}^{nf_2f_1}\frac{1}{nf_1-x}
\end{gather}
This is essentially an early termination of the summation in equation ~\ref{eq:6}, and can be expressed as follows:
\begin{gather}
    E(t) = n\sum_{x=0}^{nf_2f_1}\frac{1}{nf_1-x} = n(\sum_{x=0}^{nf_1-1}\frac{1}{nf_1-x} - \sum_{x=nf_2f_1+1}^{nf_1-1}\frac{1}{nf_1-x}) \\
    E(t) = n[H_{nf_1} - H_{nf_1 - nf_2f_1 - 1}]\\
\end{gather}
Again, using the bounds on the nth Harmonic Number~\cite{harmonicNumber}, substituting $\frac{E(t)}{64} = E(H)$ and $x$, and simplifying, we get the following: \\
\makebox[\textwidth][l]{Let}
\begin{gather}
% 	n[
%     \ln{\frac{nf_1}{nf_1 - nf_2f_1 - 1}} + \frac{1}{2nf_1} - \frac{1}{12({nf_1})^2+2\frac{7-12\gamma}{2\gamma - 1}} \\ - \frac{1}{2(nf_1 - nf_2f_1 - 1)} + \frac{1}{12({nf_1 - nf_2f_1 - 1})^2+2\frac{7-12\gamma}{2\gamma - 1}}] < E(t) \\
%     E(t) <= n(\ln{\frac{nf_1}{nf_1 - nf_2f_1 - 1}} + \frac{1}{2nf_1} - \frac{1}{12({nf_1})^2+\frac{6}{5}} \\
%     - \frac{1}{2(nf_1-nf_2f_1-1)} + \frac{1}{12({nf_1-nf_2f_1-1})^2+\frac{6}{5}}) \\
%     x  = E(H) - \frac{n}{64}\ln{\frac{nf_1}{nf_1 - nf_2f_1 - 1}} - \frac{1}{128f_1} + \frac{n}{768({nf_1})^2+128\frac{7-12\gamma}{2\gamma - 1}} \\
   \frac{n}{768({nf_1 - nf_2f_1 - 1})^2+128\frac{7-12\gamma}{2\gamma - 1}} - \frac{1}{768({nf_1})^2+128\frac{7-12\gamma}{2\gamma - 1}} < x \\
   x <= \frac{n}{768({nf_1-nf_2f_1-1})^2+\frac{6}{5}} - \frac{n}{12({nf_1})^128+\frac{6}{5}} 
\end{gather}    
%     \ln{()}) \\
%     E(t) = n\log{\frac{1}{1-f_2-\frac{1}{nf_1}}}

% Again, we divide by 64 to obtain the expected number of hash computations:
% \begin{gather}
%     E(H) = \frac{n}{64}\log{\frac{1}{1-f_2-\frac{1}{nf_1}}}
% \end{gather}
% \end{proof}
Now, using ~\ref{eq:13}, if we set $n = 16777186$, $f_1 = 1$ and $f_2 = 0.95$, we get $E(H) = 785312$ hashes

@@

Hence, we see that even computing merely 800k hashes fills a DAG buffer as large as the entire DAG by more than 95\%, as opposed to calculating 4.5 million to fill the buffer entirely. (Note that this number should be smaller for smaller buffer sizes; so this is again a rough upper bound). This suggests that the fuller the buffer is, the marginal cost of computing more hashes outweighs the rate at which the buffer is filled up. In fact, as we will see in the results, filling up 95\% of a buffer as large as the DAG already starts giving us good hash rates for a browser (10kH/s). 
This means that reaching a reasonably steady state is not as hard as it seems at face value, making this approach viable for web setting where users might not stay on websites for long. Given this result, we compute 800k hashes for all our experiments, as discussed in the next section.
% Beyond that, the marginal cost of computing more hashes outweighs the rate at which the DAG is filled up. Hence, as we will see in the next section, for our experiments, we compute 800,000 hashes.

\section{Results}
\subsection{Experimental set up}
Our experimental set up consisted of a machine with an Intel i7-7700HQ processor with 8 cores and 16GB ram. These results were obtained from a private Ethereum test network at Epoch 0. The DAG size was 16777186 slices (1.074GB). The cache size was 1.677MB. We ran the implementations in ~\verb|JavaScript| and ~\verb|WebAssembly| in the browser, and a native miner written in ~\verb|C++| that employs the same approach outside of the browser for control results. Each miner was run till 800,000 hashes were computed and the hash rate and buffer hit rate were sampled every 10,000 hashes. (Note that both the hash rate and buffer hit rate sampled at a particular time reflected the values over the 10,000 most recent hashes).
